---
title: "Bonus #1. datetime & dateutil"
subtitle: "Working with dates in Python"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2023-08-13"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

It is many times tricky to work with dates. Here I will collect and share relevant functions that I found useful along the way. 

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```


# datetime vs string

with the `datetime` package, we can easily transform between datetime/date object and its string representations.

```{python}
import datetime

date_dt = datetime.date(2023, 8, 13)
print(date_dt)
print(type(date_dt))

date = date_dt.strftime('%Y-%m-%d')
print(date)
print(type(date))

date_dt_ = datetime.datetime.strptime(date, '%Y-%m-%d')
print(date_dt_)
print(type(date_dt_))

```
# Get a list of dates

Say the task here is to get all the first day of each month from 2023-01-01 to 2023-12-01.

## method 1


```{python}
dates = []

for month in range(1,13):
  date  = datetime.date(2023, month, 1)
  dates.append(date)
  
print(dates)

```


## method 2

Alternatively, we could just use monthly incremental offered by the `relativedelta` method in the `dateutil` package. Compare to method 1, there is no hard-coded year so it could expand across years seamlessly.

```{python}
from dateutil.relativedelta import relativedelta

from_date = datetime.date(2023,1,1)
to_date = datetime.date(2023,12,1)
current_date = from_date
dates = []

while current_date <= to_date:
  dates.append(current_date)
  current_date += relativedelta(months = 1)

print(dates)

```



