---
title: "Bonus #1. complete()"
subtitle: "R to Python Translation of Cool R Functions"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-08"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

This repo was initially set up only for Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).  

But I will add some R to Python translation of loved R functions as bonus :) 

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```


# Bonus #1. complete()

## Set up

Imagine a scenario, we have some products and we hope to track their sale over months. For each product, we hope to have its sale `this month` compare to the previous month. This can be done easily using `lag()` in R or `shift()` in Python, IF our dataframe is complete.

However, our issue here is we don't have values for all the months. For example, product 1 only has sale values for month 1 and 2 (3 and 4 missing), and product 2 only has values for month 2 and 4, with 1 and 3 skipped ... (please see the hypothetical dataframe below)



```{r}
library(tidyverse) #arrange in dplyr, complete in tidyr

set.seed(123)
df <- data.frame(
  product_id = c(1:4,1:2),
  month = c(1:4,2,4),
  sale = runif(6, min=1, max=50) |> round()
)

df |> 
  arrange(product_id, month)

```


## complete() in R

Before I can apply `lag()`, I must make sure all the missing combinations (e.g., month 3 for product 1) are added to the dataframe with the sales value equals 0. This is super easy to do in R with `complete()`. As you can see now, each product now has sales values from month 1 to 4, and months with no sales value recorded now have sale value as 0.

```{r}
df |> 
  complete(product_id, month, fill=list(sale=0))
```


## To do this in Python

```{python}
df = r.df # get the df from R

df
```

To do this in Python is a bit more complex. Please see steps below, inspired by [this post on stackoverflow](https://stackoverflow.com/questions/52104737/python-how-to-expand-a-pandas-dataframes-rows-to-include-all-combinations-of). 

I will revisit this from time to time for a better way. Any comments and suggestions will be highly appreciated! You can find me on [Twitter](https://twitter.com/mena_wang), [LinkedIn](https://www.linkedin.com/in/mena-ning-wang/), and [GitHub](https://github.com/MenaWANG). 


```{python}
import pandas as pd
import itertools

df.set_index(['product_id', 'month'])\
    .reindex(pd.MultiIndex\
    .from_tuples(itertools.product(df.product_id.unique(), df.month.unique())))\
    .reset_index()\
    .rename(columns={'level_0':'product_id', 'level_1':'month'})\
    .fillna(0)
```

