---
title: "Method #27 Finding the Distribution of Values"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "`r format(Sys.Date(), '%B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame([[1, 'A'], 
                   [2, 'B'], 
                   [1, 'A']], 
                  columns = ["col1", "col2"])

print(df)

```

## Load the dataframe into R

```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #27 Finding the Distribution of Values

## Python

In Python, we can use `value_counts()` to find frequency of each unique values, and use `sort` to keep the most frequently appeared value at the top.

```{python}
df["col2"].value_counts(sort = True)

```

We can get a percentage instead of value counts by adding the `normalize = True` argument.

```{python}
df["col2"].value_counts(sort = True, normalize = True)\
          .round(3) * 100 # round it up if you like
```

## R

In R, this is easy to do with `count()`

```{r}
library(dplyr)

df |> count(col2, sort = TRUE)

```

To add percentage takes a bit more code, but you can have a nice table with both the counts and percentage in it together.

```{r}
df |> 
  group_by(col2) |> 
  summarize(n = n()) |> # get counts
  mutate(pct = n / sum(n) * 100) |> # get percentage
  arrange(n |> desc()) # sort descending

```

We could also use `tabyl()` in the `janitor` package to do it in one line. 

```{r}
library(janitor)

df |> tabyl(col2)

```

`tabyl` has much advanced functionalities that would allow us to create customized frequency tables. Please see below two simple examples and much more available [here](https://cran.r-project.org/web/packages/janitor/vignettes/tabyls.html) 

```{r table.width=10}

humans <- starwars |> filter(species == "Human")

# one-way table
humans |> 
  tabyl(eye_color) |> 
  adorn_totals() |> 
  adorn_pct_formatting(digits = 1)  # format the percentage

# two-way table
humans |> 
  tabyl(eye_color,gender) |> 
  adorn_totals(c('row','col')) |> # add  total for both rows and columns
  adorn_percentages("all") |> # pct among all for each cell
  adorn_pct_formatting(digits = 1)  |> 
  adorn_ns() # add counts 

```

