---
title: "Method #18. Filtering a DataFrame1: Boolean Filtering"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-11"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

## Create a dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame({'col1': [1,5,3],
                   'col2': [8,4,10],
                   'col3': ['A','B','B']})
df
```

## Load the dataframe into R

```{r}
df <- py$df #load the df object created in Python above

df
```


# Method #18. Filtering a DataFrame1: Boolean Filtering

## Filter with a value

A row can be selected when the condition specified is evaluated to be True for it. For example 

### Python

```{python}
df[df['col2']>5]

```


```{python}
# or to improve readability, we could do it in two steps
col2_larger_than_5 = df['col2'] > 5
df[col2_larger_than_5]

```

### R

```{r}

library(dplyr)

df |> 
  filter(col2 > 5)
```

## Filter with a list

A row can be selected when its value belongs to a list specified. For example

### Python

```{python}
filter_list = ["B","C"]

df[df['col3'].isin(filter_list)]

```

Alternatively, this can be done with `query()` in Python. `@` in the query indicating that we are using a variable in the environment.

```{python}
df.query('col3 in @filter_list')

```

### R

```{r}
filter_list = c('B','C')

df |>  
  filter(col3 %in% filter_list)
  
```

## Multiple conditions

Now let's try filter the dataframe with both conditions

### Python

```{python}
df[(df['col2']>5) & (df['col3'].isin(filter_list))]

```

For me, the syntax is probably more readable with `query()`

```{python}
df.query('(col2 > 5) and (col3 in @filter_list)')

```

### R

The syntax for dplyr is pretty similar to that with `query()` above.

```{r}
df |> 
  filter(col2 > 5 &
         col3 %in% filter_list)

```

## filter in R vs Python

One thing that might be a bit confusing for R users learning Python and vice versa is probably the use of `filter()` in these two languages. In R, `filter` is used to keep rows by certain condition, very much like `query` in python, as shown in the examples above. In contrast, `filter` is used to either select rows by indices or columns by names in Python. Below please see its four arguments, `like` is my favorite. 

- item – Takes list of axis labels that you wanted to filter.
- like – Takes axis string label that you wanted to filter
- regex – regular expression
- axis: 0 or 'index', 1 or 'columns’. When not specified it uses columns

### Python

```{python}
# first, let's create a dataframe with more complex names 
technologies= {
    'Courses':["Spark","PySpark","Spark","Java","PySpark","PHP"],
    'Fee' :[22000,25000,23000,24000,26000,27000],
    'Duration':['30days','50days','30days','60days','35days','30days']
          }
df = pd.DataFrame(technologies)
print(df)

```


```{python}
df.filter(items = ['Courses', 'Duration'])

```

I think `like` argument is very useful when we have group of columns sharing meaningful names. For example, in a large covid datset, we can use `vaccin` to filter all columns relate to `vaccination` or `vaccine`. 

```{python}
df.filter(like = 'tion')

```

We could also use filter to select rows by index, for example

```{python}
df.filter(items=[1,3], axis =0)
```

We can use query to do this as well, but it is not restricted to indices. 

```{python}
df.query("index in [1,3]")
```

```{python}
df.query("Courses in ['Spark','PySpark']")

```

`filter()` in Python and `select()` in R are also discussed in [**Method #19**](method19-filtering-a-dataframe2-getting-columns.html) on selecting columns.
