---
title: "Method #8. Modifying the Datatype of a Column"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-01"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #8. Modifying the Datatype of a Column

## Python

```{python}
import pandas as pd

data = {'id': [1,2,3,4,5],
        'name': ['Jason', 'Jackie', 'Joe', 'Joshua', 'James']}
      
df = pd.DataFrame(data)

df.dtypes
```

The id field is of integer type. Use `astype()` method to change it to string/character type as follows. 

```{python}
df2 = df.copy()
# change the id field to string/character type
df2['id'] = df2['id'].astype(str)
df2.dtypes

```


## R

```{r}
# get the dataframe from python
df = py$df

library(dplyr)

df |>  glimpse()
```

Use the `as.character()` function to change the id field into string/character type, and use `as.numeric()` function to change it back to numeric. (id: what did I do? :p)

```{r}

# change the id field to string/character type
df = df |> mutate(id = as.character(id))
df |>  glimpse()

# change the id field back to numeric type
df = df |> mutate(id = as.numeric(id))
df |> glimpse()

```


