---
title: "Method #12. Joining DataFrames"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-08"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).  

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #12. Joining DataFrames

## Python

We can use the `pd.merge()` method to join two dataframes. Join on multiple columns can be done easily.

```{python}
import pandas as pd

df1 = pd.DataFrame({'col1': [1,3,5],
                    'col2': [2,4,6],
                    'col3': ['A','A','B'],
                    'col4': ['x','y','y']})
                   
df2 = pd.DataFrame({'col3': ['A','B','C'],
                    'col4': ['x','y','z']})
                    
                    
pd.merge(df1, df2, on = ['col3','col4'], how = 'right') # the default behavior is an 'inner' join

```


In [another article by Avi](https://towardsdatascience.com/its-time-to-say-goodbye-to-the-merge-method-in-pandas-b7f39a4a3fb6), he also discussed the advantage of `join()` over `merge()` method in terms of run-time performance. Therefore, it would be a good idea to look at how `join()` works as well.  


```{python}
# set the columns to be joined on as index first
df1.set_index(['col3','col4'], inplace = True) 
df2.set_index(['col3','col4'], inplace = True)

df = df1.join(df2, how = 'right') # the default behavior is a 'left' join
print(df)

```

```{python}
df.reset_index(inplace = True)
print(df)
```


```{python}
df1.reset_index(inplace = True)
df2.reset_index(inplace = True)
```


## R

In `tidyverse`, R has separate functions for `inner_join()`, `left_join()`, etc. You specify what column(s) to join on within the function call. If we don't specify, columns with the same names will be matched by default. 

```{r}
library(tidyverse)

df1 <- py$df1
df2 <- py$df2

df <- right_join(df1, df2, by = c('col3' = 'col3', 'col4' = 'col4'))

#since we are joining on a column that has the same name in both dataframes the by= argument can be dropped
df <- right_join(df1, df2)

df
```


