---
title: "Method #30 Pivoting DataFrames"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-12-09"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.read_csv("./data/df_method30.csv")

print(df)

```

## Load the dataframe into R

```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #30 Pivoting DataFrames

## Python

In Python, `pivot_table()` can convert the column entries to column headers.

```{python}
df_wider = pd.pivot_table(df, 
               index = ["Name"],
               columns=["Subject"], 
               values='Marks',
               fill_value=0)
print(df_wider)              

```

## R

In R, this can be done with the `pivot_wider()` function. 

```{r}
library(tidyr)

df_wider = df |> 
  pivot_wider(id_cols=Name, 
              names_from=Subject, 
              values_from = Marks, 
              values_fill = 0)

print(df_wider)

```

# Bonus: Unpivoting

## R

Above we used the `pivot_wider()` function to turn a table from long to wide. Naturally, we can reverse the process and turn a table from wide to long. And the function to do this is, you guessed it, `pivot_longer()` :P

```{r}

df_wider |> 
  pivot_longer(cols = c(Maths, Science, English),
               names_to = "Subject",
               values_to = "Marks")

```

## Python

Similar task can be done by the `melt()` method in Pandas.

```{python}
df_wider.index # remember that "Name" is set as index above
df_wider.melt(ignore_index = False, # use index as the id field
              value_vars=["Maths", "Science", "English"],
              value_name = "Marks")

```

For dataframe where the index is not the id field we want to use, it is simple as well.

```{python}
df_wider.reset_index(level=["Name"], inplace=True) # reset name to be a column
df_wider.melt(id_vars = "Name", # set id field manually
              value_vars=["Maths", "Science", "English"],
              value_name = "Marks")

```

