---
title: "Method #19 Filtering a DataFrame2: Getting Columns"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-12"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in python

```{python}
import pandas as pd

df = pd.DataFrame({'col1': [1,5,3],
                   'col2': [8,4,10],
                   'col3': ['A','B','B']})
df
```

## load it in R

```{r}
df <- py$df
df
```


# Method #19. Filtering a DataFrame2: Getting Columns

## Python

For example, when we want to select col2 in the above dataframe.

```{python}
df['col2']
```

Or when you want to select multiple columns together, just specify them as a list.

```{python}
df[['col1','col3']]
```


## R

To do this in R, we can use the `select()` function

```{r}

library(dplyr)

df |>  select(col2) # select col2

df |>  select(col1, col3) # select co1 and col3
```


# Bonus: Select Columns by Conditions

Sometimes, we would want to select columns with certain characteristics

## Select columns by string in names

First let's change the column names.

```{python}
df.columns = ['apple','app','orange'] # change column names
df
```

### R

To select columns where names contain certain strings, we can simply add `contains()` function to the `select` call introduced above. 

```{r}
df = py$df #get the dataframe from python

df |> select(contains('app'))
```

### Python

To select columns where names contain certain strings, we can use the `filter()` method. 

```{python}
df.filter(like='app')
```

Special note for R users. In R, `filter()` is normally used to filter rows, whereas `select()` for columns. Here in Pandas, by default, `filter()` looks at columns of dataframes. But we can also set this to rows by specifying axis = 0

```{python}
df.set_index(['orange'], inplace=True) # set index to be the 'orange' row
print(df.index)

df.filter(like='B', axis=0) # filter df by index
```

Here index is used for filtering. To filter rows by values of variables, we can use the `query()` method in Python. See details in method [#18](method18-filteringa-dataframe1-boolean-filtering.html)


## Select columns by data types

### Python

To select columns by data types, we can use the `select_dtypes` method in python

```{python}
df.dtypes #check data types
df.select_dtypes(include = ['int64'])
df.select_dtypes(exclude = ['int64'])
```

To use this to identity all numeric fields, we can set a more general criteria like below

```{python}
import numpy as np
num_features = df.select_dtypes(include = np.number).columns.tolist()
print(num_features)
```


### R

To do this in R, simply add a `where()` in the select call introduced above.

```{r}

str(df)#check data types
df |> select(where(is.numeric))
df |> select(-where(is.numeric))
```



