---
title: "Method #25-26 Handling Duplicates"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-25"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```


# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in Python

In the simple dataframe below, the 1st and 3rd row are the same.

```{python}
import pandas as pd

df = pd.DataFrame([[1, 'A'], 
                   [2, 'B'], 
                   [1, 'A']], 
                  columns = ["col1", "col2"])

print(df)

```

## Load the dataframe into R

```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #25-26. Handling Duplicates

## Python

In Python, we can use `duplicated()` to identify duplicated rows. 

```{python}
df.duplicated() #the default setting only marks the second appearance of the duplicated rows
df.duplicated(keep=False) #add `keep = False` to identify all appearances of the duplicated rows
```

The above result can be used to identify duplicated rows in a dataframe. Check out how the results below are slightly different. 

```{python}
df[df.duplicated(keep=False)]
```

```{python}
df[df.duplicated()]
```

To drop duplicated rows in Python, we can use the `drop_duplicates()` method as follows. 

```{python}
df.drop_duplicates()
```

## R

In base R, we also have a `duplciated()` function, notice that only one of the row `[1, 'A']` was marked as duplicated.  

```{r}
df |> duplicated()

# use the result to identify duplicated rows
df[duplicated(df),]

```

To drop duplicated rows in R, we can use the `unique()` method in base R.

```{r}
df |> unique()

```


