---
title: "Method #10. Printing Descriptive Info about the DataFrame (Method 2)"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-03"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).  

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #10. Printing Descriptive Info about the DataFrame (Method 2)

## Python

`describe()` returns standard statistics like mean, standard deviation, maximum etc. of every numeric-valued column

```{python}
import pandas as pd

df = pd.read_csv('data/iris.csv')

df.describe()
```

`describe()` can also offer some info on categorical columns: the number of unique values, the most frequent value and its frequency, if we add `include = all` argument
```{python}
df.describe(include = 'all')
```


## R

As discussed in Method 9, `summary()` in base R returns the number of missings as well as some summary statistics for all numerical and factor columns.  

```{r}

df = py$df

df |>  summary()

```


For categorical variables, like `Species` in the iris dataset, `summary()` could give us its frequency counts if we turn it into factor type in R. 

```{r}

library(dplyr) #for mutate

df |> 
  mutate(Species = as.factor(Species)) |> # see method17 for more on mutate()   
  summary()

```

In python, we could use `value_counts()` to get frequency counts. See #method27 for more.

```{python}
df['Species'].value_counts()

```



