---
title: "Method #6. Viewing Top N Rows"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-01"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #6. Viewing Top N Rows

## Python

Use `head()` to show the top N rows. 

```{python}
import pandas as pd

df = pd.read_csv('data/iris.csv')

# print the top 3 rows in the dataframe
print(df.head(3))

```


We can also sort the dataframe first then show the top N. For more on sorting a dataframe, please refer to method 13.

```{python}
# to sort the dataframe first then check out the top N
df.sort_values('Sepal.Length')\
  .head(5)

```


## R

To do these tasks in R is pretty straightforward. 

```{r}

df <- py$df # get the dataframe from python

# print the top 3 rows in the dataframe
print(df |>  head(3))

```

To sort first, we can use the `arrange()` function in the dplyr package.

```{r}

library(dplyr) 

# to sort the dataframe first then check out the top N
df |> 
  arrange(Sepal.Length) |> 
  head(5)

```


