---
title: "Method #24 Apply A Function to A Dataframe"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-18"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame([[1, 2], 
                   [5, 8], 
                   [3, 9]], 
                  columns = ["col1", "col2"])

print(df)

```

## Load the dataframe into R


```{r}
df <- py$df #access df as saved in Python(py) above

df <- data.frame(
  col1 = c(1,5,3),
  col2 = c(2,8,9)
)

print(df)
```


# Method #24. Apply A Function to A Dataframe

## Python

In Python, we can use `apply()` to apply a function to a dataframe. Let's define a simple function first.

```{python}
# create a function to be applied
def add_cols(row):
  return row.col1 + row.col2

```

Let's apply the function to all the columns in a datafarme

```{python}
df['col3'] = df.apply(add_cols, axis = 1)
df
```

Or we could only apply a function to one particular columns in a dataframe.

```{python}
def square_col(num):
  return num**2

df['col3']=df.col2.apply(square_col)
df
```


## R

In R, we could use `lapply()` in dplyr package to apply a custom function to all values in a dataframe, note that the results are in multiple lists rather than one dataframe. 

```{r}

library(dplyr)

square_col = function(num){
  return (num^2)
}

df2 = df |> lapply(FUN=square_col)
df2 |> str() #lapply returns result in lists
df2 |> data.frame() |> str() # transform the lists into a dataframe

```


We can use the `map_df()` function in `purrr()` package to do the similar task but generate a dataframe (tibble) directly. What is a tibble? You might be wondering. well, think of it basically as a dataframe, just even nicer. Read more about it [here](http://statseducation.com/Introduction-to-R/modules/getting%20data/tibbles/) if interested.  

```{r}
library(purrr)

df |> map_df(square_col) 

```

To apply a function to one column to create a new column without changing the existing column itself, the suitable mapping method depends on the type of output. When the output is a numeric column, use `map_dbl()`, whereas when the output is a character column, use `map_chr()`. Below please see an example for `map_dbl()`, and more illustrations are available [here](https://www.geeksforgeeks.org/apply-function-to-every-value-in-r-dataframe/#:~:text=In%20R%20Programming%20Language%20to,use%20paste()%20with%20lapply.). For other ways to create new columns in Pandas and R, please refer to method 17.


```{r}

df['col3'] = df[['col2']] |> map_dbl(square_col)
df

```

In contrast, if we want to apply a function to change a subset of columns in a dataframe, we could simply apply `select()` function before mapping the function to the dataframe.

```{r}

df |> select(col2) |> map_df(square_col)
```





