---
title: "Method #28 Resetting the Index of a DataFrame"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-12-02"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

## Create the dataframe in Python

Notice the index is set to be 2,3,1

```{python}
import pandas as pd

df = pd.DataFrame([[6, 5,  10], 
                   [5, 8,  6], 
                   [3, 10, 4]], 
                  columns = ["col1", "col2", "col3"],
                  index = [2, 3, 1])

df.head()

```

## Load the dataframe into R

Notice that the same index has been passed on to R

```{r}
df <- py$df #access df as saved in Python(py) above

df |> head()

```


# Method #28 Resetting the Index of a DataFrame

## Python

In Python, we can use the `reset_index()` method to reset the index. But before that, let's quickly review how we can choose values by index or by position (method [#20](method20-filtering-a-dataframe3-select-by-label.html) and [#21](method21-filtering-a-dataframe4-select-by-position.html)).


```{python}
df.loc[2] # we can use loc to get values with index
df.iloc[2] # we can use iloc to get values by position

df.reset_index(drop = True) #drop the old index
df
```

A bit more play with `loc[]` and `iloc[]` based on the newly set index

```{python}
df.loc[1:2] # loc to get values with index
df.iloc[[1,2],] # iloc to get values by position
```

## R

In R, we can get values by positions using `slice()`

```{r}
library(dplyr)

df |> slice(1:2) # choose the 1st and 2nd row

```

To choose rows by name, we need to tell R the row name using a list like this.

```{r}

df[c('1','2'),]

```

To reset the row names in R

```{r}

rownames(df) = 1:nrow(df)
df

```

