---
title: "Method #15. Renaming Column(s)"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-09"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/). 


# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #15. Renaming Column(s)

## Python

We can use the `rename()`` method in Pandas to rename a column in a dataframe. 

```{python}
import pandas as pd

df = pd.DataFrame([[1, 2,  "A"], 
                   [5, 8,  "B"], 
                   [3, 10, "B"]], 
                  columns = ["col1", "col2", "col3"])

df.rename(columns = {'col1': 'col1_new_name', 'col2': 'col2_new_name'}, inplace = False) # this doesn't change df unless we set inplace to be True

df

```

## R

In R there is a similar function `rename()`, just note that the sequance is `new_name = old_name`, not `old_name : new_name` as shown above in Python. 

```{r}

library(dplyr)

df <- py$df

df |>  
  rename(col1_new_name = col1,
         col2_new_name = col2)

```

