---
title: "Method #20 Filtering a DataFrame3: Selecting by Label (.loc)"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "`r format(Sys.Date(), '%B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame([[6, 5,  10], 
                   [5, 8,  6], 
                   [3, 10, 4]], 
                  columns = ["Maths", "Science", "English"],
                  index = ["John", "Mark", "Peter"])

print(df)

```

## Load the dataframe into R

We notice that the index has been passed on to the dataframe in R.

```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #20. Filtering a DataFrame3: Selecting by Label

## Python

with `.loc[]`, we can select by labels as follows

```{python}
df.loc['John'] # select one specific row

df.loc['John','Maths'] # select one specific row and column

df.loc[['John','Mark'],['Maths','Science']] # select a list of rows and columns
```

## R

To do this in base R is fairly simple as shown below

```{r}

df['John',] # select one specific row

df['John','Maths'] # select one specific row and column

df[c('John','Mark'),c('Maths','Science')]  # select a list of rows and columns
  
```

# Bonus: select columns by value

Imagine a dataframe like below, we have scores of multiple students and would like to identify all students who score 80 or above in all three courses.

```{python}

data = {
  'subject': ['math', 'science', 'english'],
  'John': [80, 95, 78],
  'Alex': [90, 67, 84],
  'Angela': [87, 86, 90],
  'Bella': [78, 91, 80],
  'Trice': [87, 75, 91]
}

df = pd.DataFrame(data)

df.set_index('subject', inplace = True)

df

```


## python

In python, `loc` can be used to identify students who score higher than 80 in each course. 

```{python}

math_80 = df.loc[:, df.loc['math']>80].columns
science_80 = df.loc[:, df.loc['science']>80].columns
english_80 = df.loc[:, df.loc['english']>80].columns
print(f'{math_80}\n{science_80}\n{english_80}')

```

We can turn the lists into set then use `set.intersection` to find common elements in it. 

```{python}
all_rounder = list(set.intersection(*map(set, [math_80, science_80, english_80])))
all_rounder
```


## R

To do this in r, `select_if` which is a variation/extension of `select` introduced in [method #19](method19-filtering-a-dataframe2-getting-columns.html) can be used to identify students who score above 80 in each course.

```{r}
library(dplyr)
df = py$df

math_80 <- df |> select_if(df['math',] > 80) |> colnames()
science_80 <- df |> select_if(df['science',] > 80) |> colnames()
english_80 <- df |> select_if(df['english',] > 80) |> colnames()

```

Then use `intersect` to identify common elements.

```{r}
all_rounder = Reduce(intersect, list(math_80, science_80, english_80))
all_rounder
```

