---
title: "Method #7. Printing the Datatype of Dolumns"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-01"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #7. Printing the Datatype of Columns

## Python

```{python}
import pandas as pd

df = pd.read_csv('data/iris.csv')

# .dtypes returns the data type of all columns
df.dtypes

```


## R

The `str()` function in base R returns the data type of all columns, the default option also returns the dimensionality of the dataframe, length and head of each column, and gives attributes as sub structures. 

```{r}
#get the dataframe from python
df = py$df

df |> str(give.attr = FALSE)
```

Another option is `glimpse()` function in the dplyr package, which returns the dimensionality of the dataframe, data type of all columns and the head of each column as well.

```{r}

library(dplyr)

df |> glimpse()

```


