---
title: "Method #19 Filtering a DataFrame2: Getting Columns"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-12"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in python

```{python}
import pandas as pd

df = pd.DataFrame({'col1': [1,5,3],
                   'col2': [8,4,10],
                   'col3': ['A','B','B']})
df
```

## load it in R

```{r}
df <- py$df
df
```


# Method #19. Filtering a DataFrame2: Getting Columns

## Python

For example, when we want to select col2 in the above dataframe.

```{python}
df['col2']
```

Or when you want to select multiple columns together, just specify them as a list.

```{python}
df[['col1','col3']]
```


## R

To do this in R, we can use the `select()` function

```{r}

library(dplyr)

df |>  select(col2) # select col2

df |>  select(col1, col3) # select co1 and col3
```


