---
title: "Method #8. Modifying the Datatype of a Column"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2023-07-21"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #8. Modifying the Datatype of a Column

## Python

```{python}
import pandas as pd

data = {'id': [1,2,3,4,5],
        'name': ['Jason', 'Jackie', 'Joe', 'Joshua', 'James']}
      
df = pd.DataFrame(data)

df.dtypes
```

The id field is of integer type. Use `astype()` method to change it to string/character type as follows. 

```{python}
df2 = df.copy()
# change the id field to string/character type
df2['id'] = df2['id'].astype(str)
df2.dtypes

```
Sometimes we might have multiple columns to be transformed, this can be done super easily with `astype()`.

```{python}
# Sample DataFrame
data = {
    'col1': [1, 2, 3],
    'col2': [4, 5, 6],
    'col3': [7, 8, 9]
}

df3 = pd.DataFrame(data)

# Convert specific columns from numeric to string
columns_to_convert = ['col1', 'col2', 'col3']
df3[columns_to_convert] = df3[columns_to_convert].astype(str)

# Verify the data types
print(df3.dtypes)

```
To change strings to numeric type, we could use `astype(int)`, but this will throw an error when encountering non-numeric characters. Therefore, I prefer using `pd.to_numeric()` instead, which offers the `errors` parameter to determine how we want to handle non-numeric characters.  

```{python}
# Sample DataFrame
data = {
    'col1': ['1', 'Unknown', '3'],
    'col2': ['4', 'N.A.', '6'],
    'col3': ['7', '8', 'Missing']
}

df4 = pd.DataFrame(data)

# Using 'coerce': Non-numeric values will be converted to NaN
df4['col1'] = df4['col1'].apply(pd.to_numeric, errors='coerce')

# Using 'ignore': Non-numeric values will be left unchanged
df4['col2'] = df4['col2'].apply(pd.to_numeric, errors = 'ignore')

print(df4)
print(df4.dtypes)

```
```{python}
# Using 'raise': This will raise an error due to the presence of non-numeric values
try:
    df4['col3'] = pd.to_numeric(df4['col3'], errors='raise')
except ValueError as e:
    print("Error:", e)
```



## R

```{r}
# get the dataframe from python
df = py$df

library(dplyr)

df |>  glimpse()
```

Use the `as.character()` function to change the id field into string/character type, and use `as.numeric()` function to change it back to numeric. (id: what did I do? :p)

```{r}

# change the id field to string/character type
df = df |> mutate(id = as.character(id))
df |>  glimpse()

# change the id field back to numeric type
df = df |> mutate(id = as.numeric(id))
df |> glimpse()

```


