---
title: "Method #22-23 Finding Unique Values in a DataFrame"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-17"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame([[6, 5,  10], 
                   [5, 8,  6], 
                   [5, 10, 6], 
                   [4, 7,  10]],
                  columns = ["Maths", "Science", "English"])

print(df)

```

## Load the dataframe into R


```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #22-23. Finding Unique Values in a DataFrame

## Python

use `unique()` to find unique values, and `nunique()` for the number of unique values

```{python}
df['Maths'].unique() # get unique values in the 'Maths' column

df['English'].unique() # get unique values in the 'English' column

df['Maths'].nunique() # get the number of unique values in the 'Maths' column

df['English'].nunique() # get the number of unique values in the 'English' column

```

## R

To get unique values of a column, we select the column then apply the `unique()` function. This can also be used to find unique value combinations across multiple variables. To count the number of unique values or value combinations, just apply `length()` function to count its length.     

For python, `unique()` can't be used to get unique value combinations across multiple variables. If this is what you seek to do, please refer to method 26 in [Avi's article](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) which we will get to translate soon.

```{r}
library(dplyr)

# unique values in one single column
df |> 
  select(Maths) |> 
  unique()

# unique combination of values across multiple columns
df |> 
  select(Maths, English) |> 
  unique()

# count the number of unique values or value combination
df |> 
  select(Maths, English) |> 
  unique() |> 
  length()

```