---
title: "Method #29 Finding Cross-Tabulation"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-12-02"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame([["A", "X"], 
                   ["B", "Y"], 
                   ["C", "X"],
                   ["A", "X"]], 
                  columns = ["col1", "col2"])

print(df)

```

## Load the dataframe into R

```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #29 Fiding Cross-Tabulation

## Python

In Python, `crosstab()` can return the frequency of each combination of values across two columns.

```{python}
pd.crosstab(df.col1, df.col2)

```

## R

In R, `table()` method can do the similar task.

```{r}
table(df$col1, df$col2)

```

But with `tabyl()` function in the `janitor` package, we can have much more advanced functionalities and customized rosstable, more in our previous discussion on method [#27](method27-finding-the-distribution-of-values.html)
