---
title: "Method #11. Filling Missing Values"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-05"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).  

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```


# Method #11. Filling NaN values

## Python

We can use the `df.fillna()` method to replace missing values with a specific value. Let's start by creating a dataframe with missing values. 

```{python}
import pandas as pd
import numpy as np

df = pd.DataFrame({'col1': [1,2],
                   'col2': [3,np.nan],
                   'col3': ['A',np.nan]})
print(df)
```

To replace the missing value with `0`, just do `.fillna(0)`. 

```{python}
# replace all NA values with 0, inplace = True means df itself will be modefied
df.fillna(0, inplace = True)
print(df)
```


## R

We can use either `replace()` in base R to replace all missings in a dataframe with a specific value, or `replace_na` in tidyr to offer tailored replacement for each specific column.

```{r}

df = data.frame(col1 = c(1,2),
                col2 = c(3, NA),
                col3 = c('A',NA))

print(df)

# use replace() in base R to replace all missings to 0
replace(df, is.na(df), 0)

library(tidyr)
# or use replace_na() in tidyverse to offer tailored replacement for each column. 
df |>  replace_na(list(col2=0,col3="Unknown"))

```


