---
title: "Method #21 Filtering a DataFrame4: Selecting by Position"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-07"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame([[6, 5,  10], 
                   [5, 8,  6], 
                   [3, 10, 4], 
                   [4, 7,  9]],
                  columns = ["Maths", "Science", "English"])

print(df)

```

## Load the dataframe into R


```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #21. Filtering a DataFrame4: Selecting by Position

## Python

with `.iloc[]`, we can select a row by position

```{python}
df.iloc[0] # select the first row

df.iloc[0:2] # select the first two rows

df.iloc[-1] # select the last row

```

## R

To do this in R is fairly simple as shown below, also we could `slice()`, `slice_head()` and `slice_tail()` in the `dplyr` package as fit. 

```{r}
library(dplyr)

df |> slice(1) # select the first row

df |> slice(1:2) # select the first two rows

df |> slice_tail(n=1) # select the last row

df |> slice_tail(prop = 0.5) # select the bottom half

df |> slice_head(prop = 0.25) # select the top 1/4

```


To get top or bottom perc% of rows in Python, I didn't find a build-in method yet. But it could be done easily with some simple calculation

```{python}
half_rows = int(round(0.5*len(df),0))# calculate 50% of rows

df.iloc[0:half_rows] # get the top half

```

# Combining position with label identifiers?

In this note we talked about `iloc` which relies on position, while in the previous one we talked about `loc` which relies on labels. But sometimes, we might want to combine the two identifiers. Say, we want to identify the value of the 'English' column in the 1st row. 

```{python}
print('before', df.iloc[0]['English'])
df.iloc[0]['English'] = 15 # this worked here, but in some python env it doesn't ... interesting!
print('after', df.iloc[0]['English'])

```

Note that using `.iloc[]` or `.loc[]` alone, there is no issues at all.

### `iloc` alone

```{python}
print('before', df.iloc[0,2])
df.iloc[0,2] = 12
print('after', df.iloc[0,2])

```

### `loc`. alone

Note that in this df, the row index is consistent with its location, but this is not always the case. Indexing is a powerful tool in python, pls see [more here](https://github.com/MenaWANG/ML_toy_examples/blob/main/Others/test%20index.ipynb)

```{python}
print('before', df.loc[0,'English'])
df.loc[0,'English'] = 17
print('after', df.loc[0,'English'])

```

### Assigning values by both location and label

Alright, go back to the question before, how do we assign value to the 'English' column in the 1st row? One easy way is translating the label to position (or vice versa). Pls see below

```{python}
print('before', df.iloc[0]['English'])
print('location of the "English" column:', df.columns.get_loc('English'))
df.iloc[0, df.columns.get_loc('English')] = 100
print('after', df.iloc[0]['English'])

```

