---
title: "Method #21 Filtering a DataFrame4: Selecting by Position"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-07"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Provide a Python to R translation of **30 essential Pandas methods** introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)
```

## Create the dataframe in Python

```{python}
import pandas as pd

df = pd.DataFrame([[6, 5,  10], 
                   [5, 8,  6], 
                   [3, 10, 4], 
                   [4, 7,  9]],
                  columns = ["Maths", "Science", "English"])

print(df)

```

## Load the dataframe into R


```{r}
df <- py$df #access df as saved in Python(py) above

print(df)
```


# Method #21. Filtering a DataFrame4: Selecting by Position

## Python

with `.iloc[]`, we can select a row by position

```{python}
df.iloc[0] # select the first row

df.iloc[0:2] # select the first two rows

df.iloc[-1] # select the last row

```

## R

To do this in R is fairly simple as shown below, also we could `slice()`, `slice_head()` and `slice_tail()` in the `dplyr` package as fit. 

```{r}
library(dplyr)

df |> slice(1) # select the first row

df |> slice(1:2) # select the first two rows

df |> slice_tail(n=1) # select the last row

df |> slice_tail(prop = 0.5) # select the bottom half

df |> slice_head(prop = 0.25) # select the top 1/4

```


To get top or bottom perc% of rows in Python, I didn't find a build-in method yet. But it could be done easily with some simple calculation

```{python}
half_rows = int(round(0.5*len(df),0))# calculate 50% of rows

df.iloc[0:half_rows] # get the top half

```

