---
title: "Method #16-17. Adding New Column(s)"
subtitle: "Python to R Translation of 30 Essential Pandas Methods"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2022-11-10"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

```

# Introduction

Python to R translation of 30 essential Pandas methods introduced by Avi Chawla in [The Only 30 Methods You Should Master To Become A Pandas Pro](https://towardsdatascience.com/the-only-30-methods-you-should-master-to-become-a-pandas-pro-749795084bb2) published on [TowardsDataScience](https://towardsdatascience.com/).

# Set up

```{r}
# enable python in RMarkdown
library(reticulate)

```

# Method #16-17. Adding New Column(s)

## Python

To add a column based on simple calculation of existing columns, we could either do the calculation directly (first chunk below) or use the `assign()` method. 


```{python}
import pandas as pd

df = pd.DataFrame({'col1': [1,2],
                   'col2': [3,4]})
```

```{python}
# direct calculation
df2 = df.copy()
df2['col3'] = df['col1'] + df['col2'] 
df2
```


```{python}
# use assign() method
df2 = df.assign(col3 = df['col1'] + df['col2'] )
df2
```

## R

In R, `mutate()` function handels these tasks with ease

```{r}

library(dplyr)

df <- py$df #load the df object created in Python above

df2 <- df |>  
  mutate(col3 = col1 + col2)
df2
```


# Bonus: More complex situations

Now let's try some slightly more complex calculations and/or create more than one columns in one go as well. 

## Create multiple columns in one go

### Python

```{python}
(df
  .assign(col3 = lambda x: x.col1*2 + x.col2)
  .assign(col4 = lambda x: (x.col1/x.col2 * 100).round(3)))

```

### R

To accomplish a similar task in R, we will continue to use `mutate()`

```{r}
df |>  
  mutate(col3 = col1*2 + col2,
         col4 = round(col1/col2 * 100, 3))
```

## Mean across columns

### Python

```{python}
df['row_mean'] = df.iloc[:,0:].mean(axis=1) # all columns
print(df)
```

```{python}
df['row_mean'] = df.iloc[:,0:2].mean(axis=1) # specify columns by range
print(df)
```

```{python}
df['row_mean'] = df.iloc[:,[0,1]].mean(axis=1) # specify columns by position
print(df)
```

### R

```{r}
df |> mutate(row_mean = df |> rowMeans()) # all columns
df |> mutate(row_mean = df |> select(1:2) |> rowMeans()) # specify columns by range
df |> mutate(row_mean = df |> select(1,2) |> rowMeans()) # specify columns by position

```

## Encoding categorical columns

### Python

The `map` function is pretty helpful in the scenario

```{python}
df = pd.DataFrame({'gender':['male', 'female', 'male', 'female']})
# add a column with gender encoding
df['is_male'] = df['gender'].map({'male': 1, 'female': 0})
print(df)

```

### R

Of course, tihs task is easy with dplyr

```{r}
df <- py$df
df |> 
  mutate(is_male = ifelse(gender == 'male', 1, 0))
```





