---
title: "SQL #1 SPLIT_PART"
subtitle: "Collection of SQL Tips"
author: "[Mena WANG](https://www.linkedin.com/in/mena-ning-wang/)"
date: "2023-10-07"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{css, echo=FALSE}

.python{
  background-color : Azure; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}

.r{
  background-color : LavenderBlush; #color names: https://en.wikipedia.org/wiki/Web_colors
  font-weight : bold;
}


```


# Set up

Use `RSQLite` library allows us to run SQL queries in R environment. We also have the option to create an in-memory database for quick experimentation.

```{r}
# Load the RSQLite library to use SQL in R
library(RSQLite)

# Create a SQLite database
# con <- dbConnect(SQLite(), dbname = "mydatabase.db")

# Create an SQLite in-memory database connection
con <- dbConnect(SQLite(), dbname = ":memory:")

```

Sometimes we might want to use SQL query to extract specific element of a field. On platforms such as Snowflake, I found `SPLIT_PART` very helpful in such scenarios. For example, in the query below, `SPLIT_PART` is used to extract the domain name from a URL by first splitting the URL using '://' as a delimiter and then further splitting the result using '/' as a delimiter to obtain the domain name. 

But the `SPLIT_PART` function is not supported in RSQLite, so here in the demo I will show both the SPLIT_PART way and the alternative that works in RSQLite.  

```{r}

# Create a dummy table with sample data
dbWriteTable(con, "url_demo", data.frame(
  id = 1:4,
  url = c(
    'https://www.example.com/page1',
    'https://www.example.com/page2',
    'https://www.openai.com/research/',
    'https://www.github.com/user/repo'
  )
))

# Query that would have worked in Snowflake
query <- "
SELECT
    id,
    url,
    SPLIT_PART(SPLIT_PART(url, '://', 2), '/', 1) AS domain_name
FROM
    url_demo;
"

# Define the SQL query to extract domain names
query <- "
SELECT
    id,
    url,
    SUBSTR(url, INSTR(url, '://') + 3, INSTR(SUBSTR(url, INSTR(url, '://') + 3), '/') - 1) AS domain_name
FROM
    url_demo;
"

# Execute the SQL query
result <- dbGetQuery(con, query)

# Print the result
print(result)

```


